MS Anxiety Longitudinal

### 2024-07-25 scripts/Anxiety_longitudinal_started_20240717.Rmd
Add files via upload

---
title: "MS Prelim Data K23"
author: "Erica Baller"
date: "2/25/2022"
output:
  html_document: default
  pdf_document: default
---

@@

### testing mean UF in depression group  - NS!!!
(previous script (~/BBL/msdepression/scripts/Biological_Psychiatry_Revision_Scripts/msdepression_n380_started_20230927_mm_BP_r1.Rmd)), lines 734 and  779-781
  #bilat: depGroupVar        85.32     206.70   t=0.413    p=0.680    
  disease_mean_UF_vol_lm <- lm(mean_UF_vol ~ depGroupVar + osex+PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
  summary(disease_mean_UF_vol_lm)
#findings

@@

#get list of unique patients, n = 890
mimosa_qc_unique <- mimosa_qc_df %>%
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup #n=890

#get list of unique patients with good QC, n = 867
mimosa_unique_score_gre_75 <- mimosa_qc_unique %>%
  filter(score >= 75) %>%
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  mutate(EXAM_DATA = as.character(EXAM_DATE)) %>%
  mutate(EMPI = as.character(EMPI)) %>%
  dplyr::select(EMPI, EXAM_DATE, score) %>%
  slice(1) %>%
  ungroup 
```

@@

#n=867, individual scans
df_people_with_good_mimosa <- data_2023_pull %>% 
  group_by(EMPI, EXAM_DATE) %>%
  slice(1) %>%
  left_join(mimosa_unique_score_gre_75, by = c("EMPI", "EXAM_DATE")) %>%
  filter(!is.na(score))
```

@@

 #n = 812
df_demo_and_fascicles <- df_people_with_good_mimosa %>% 
  left_join(fascicle_proportions, by = c("EMPI", "EXAM_DATE")) %>%
  filter(!is.na(Has.MS.Provider)) %>% #remove people who don't have MS provider
  filter(!is.na(CNII_L)) #this is a sample fascicle that should have a value. if it does not, it means something is usually wrong with whole person. Remove rows with NAs

@@

## Merge with data frame with good mimosas
```{r merge_lesion_volume_w_mimosa_df}
##########################################################
#             Merge lesion volumes with mimosa df        #
##########################################################
#merge whole data set with fascicle proportions, n = 812
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  left_join(lesion_volumes, by = c("EMPI", "EXAM_DATE"))

@@

df_w_yeo_volume_info <- df_demo_and_fascicles %>%
  # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  dplyr::select(EMPI, ends_with("_vol")) %>%
  ungroup()

@@

### Define anxiety net by top 25 in both vol/prop
```{r defineanxietynet_with_25_propx25vol}
##### make an array of fascicles that have both substantial overlap with the anxiety network (top 25% vol), and where each fascicle is substantially impacted (top 25% proportion), n = 14
fascicle_names_top_25_vol_x_top_25_prop <- fascicle_names_top_25_vol[fascicle_names_top_25_vol %in% fascicle_names_top_25_prop]

#add vol/prop suffixes
fascicle_names_top_25_vol_x_top_25_prop_PROP_suffix <-fascicle_names_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)
fascicle_names_top_25_vol_x_top_25_prop_VOL_suffix <-fascicle_names_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

#create fascicle vectors that don't include these special fascicles, n = 63
fascicle_names_NOT_top_25_vol_x_top_25_prop <- fascicle_names_no_cranial_nerves[!(fascicle_names_no_cranial_nerves %in% fascicle_names_top_25_vol_x_top_25_prop)]

#add vol/prop suffixes
fascicle_names_NOT_top_25_vol_x_top_25_prop_PROP_suffix <- fascicle_names_NOT_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)
fascicle_names_NOT_top_25_vol_x_top_25_prop_VOL_suffix <- fascicle_names_NOT_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

#anxiety25_propxvol overlap total, sum of these two is 1545449 (and this is total_fascicle_volume)
anxiety_top_25_vol_x_top_25_prop_vol <- sum(dep_net_overlap_from_bp_paper$num_voxels_total[dep_net_overlap_from_bp_paper$fascicle %in% fascicle_names_top_25_vol_x_top_25_prop]) #vol = 181869

#nonanxiety25_propxvol overlap total
anxiety_NOT_top_25_vol_x_top_25_prop_vol <- sum(dep_net_overlap_from_bp_paper$num_voxels_total[dep_net_overlap_from_bp_paper$fascicle %in% fascicle_names_NOT_top_25_vol_x_top_25_prop]) #vol = 1363580

#check sums
if ((anxiety_top_25_vol_x_top_25_prop_vol + anxiety_NOT_top_25_vol_x_top_25_prop_vol) == total_fascicle_volume) {
  print("anxiety + nonanxiety net = total fascicle volume")
} else{
  print("something is wrong with counts")
}

@@

    ##sum of total volume lost in anxiety 25 volxanxiety25prop groups, and proportion of volume lost in network (volume lost/network volume size)
  mutate(sum_fascicle_names_top_25_vol_x_top_25_prop = 
           rowSums(across(all_of(fascicle_names_top_25_vol_x_top_25_prop_VOL_suffix)))) %>%
  mutate(proportion_fascicle_vol_lost_top_25_vol_x_top_25_prop = sum_fascicle_names_top_25_vol_x_top_25_prop/anxiety_top_25_vol_x_top_25_prop_vol) %>%

    ##sum of total volume lost in remaining 63 fascicles NOT in anxiety 25 volxanxiety25prop groups, and proportion of volume lost in network (volume lost/network volume size)
  mutate(sum_fascicle_names_NOT_top_25_vol_x_top_25_prop = 
           rowSums(across(all_of(fascicle_names_NOT_top_25_vol_x_top_25_prop_VOL_suffix)))) %>%
  mutate(proportion_fascicle_vol_lost_NOT_top_25_vol_x_top_25_prop = sum_fascicle_names_NOT_top_25_vol_x_top_25_prop/anxiety_NOT_top_25_vol_x_top_25_prop_vol) %>%

   ##mean_uncinate)
  mutate(mean_UF_vol = (UF_L_vol + UF_R_vol)/2) %>%
  mutate(mean_UF_prop = (UF_L_prop + UF_R_prop)/2) %>%

@@

@@

### BEST SPLIT IS STRATIFYING ON ANXIETY MEDS!!!! Better than on variable that includes anxietyGroupVar(meds + anxiety diagnoses), or anxiety diagnosis alone
make_demographics_table_ms_anxiety(data_frame = df_demo_and_fascicles)
make_demographics_table_ms_anxiety(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$anxietyGroupVar != 0,])
make_demographics_table_ms_anxiety(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$Has.Anxietydx.AND.AntianxietymedGroupVar != 0,])

@@

#COLLAPSE anxiety dx and antianxiety meds with those plus depression, and anxiety dx OR on anxiety meds into 1 group -->> THIS IS THE MONEY
### Overlap with depression group
```{r overlap_with_dep_group}
#### 53.4% overlap total

#read csv of empis from depression paper
depression_empis <- read.table(file="/Users/eballer/BBL/msanxiety/data/empis_of_people_in_BP_depression_paper.csv")
names(depression_empis) <- "EMPI"

#get final anxiety dataset and make column if the empi is in the prev list
df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  mutate(EMPI = as.numeric(EMPI)) %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  mutate(EMPI_in_dep_group = EMPI %in% depression_empis$EMPI)

#"Percent overlap 53.4%"
print(paste0("Percent overlap ", round(sum(df_disease_burden_unc_and_anx_net_collapsed_variables$EMPI_in_dep_group)/length(df_disease_burden_unc_and_anx_net_collapsed_variables$EMPI), 3) * 100, "%"))

 tibble_overall_counts <- df_disease_burden_unc_and_anx_net_collapsed_variables %>% 
  group_by(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  summarize(n(),
            num_in_dep_group = sum(EMPI_in_dep_group),
            percent_overlap = num_in_dep_group/n())
 tibble_overall_counts

### Overall Disease by Anxiety Dose *acnp
```{r group_anxietydx_meds_and_dep_comorbidity}
#this collapses anxiety group + meds with those with anxiety group + meds + with/without depression

@@

  #gam, anxiety_dose  , anxiety_dose  0.03505    0.01340   2.616  0.00927 ** 
   anxietyGrouping_nonanxiety_net_gam<- gam(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_nonanxiety_net_gam)

  #volume lost in anxiety net by anxiety dose*sex: anxiety_dose         0.0459425  0.0164036   2.801  0.00537 ** 
  anxietyGrouping_by_sex_nonanxiety_net_lm <- lm(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose*osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_by_sex_nonanxiety_net_lm)

  #gam, anxiety_dose, anxiety_dose         0.047100   0.016230   2.902  0.00393 **          
  anxietyGrouping_by_sex_nonanxiety_net_gam <- gam(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose*osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_by_sex_nonanxiety_net_gam)

  #volume lost in anxiety net by anxiety dose + sex + PAT_AGE_AT_EXAM: anxiety_dose     0.0344065  0.0135309   2.543   0.0114 *  
  anxietyGrouping_cov_sex_nonanxiety_net_lm <- lm(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_cov_sex_nonanxiety_net_lm)

  #gam, anxiety_dose - anxiety_dose  0.03600    0.01342   2.682  0.00764 ** 
   anxietyGrouping_cov_sex_nonanxiety_net_gam <- gam(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_cov_sex_nonanxiety_net_gam)

@@

###Check for polynomial relationships 
```{r polynomial_relationships}
 #Not finding them w healthy, anxietydx OR meds, anxiety AND meds

 #check for polynomial relationship
#define group
df_demo_and_fascicles_no_unclass_anxiety_dose_factor <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != 3) %>% 
  mutate(anxiety_dose = as.factor(anxiety_dose))

@@

### Lesion volumes 
#### Diagnosis 
  - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, volume_of_mimosa_lesions) %>%
  ungroup()

##### healthy vs sick

#lm lesion volumes; DiagnosisMS+Anxiety  6310.99    2143.87   2.944 0.003694 ** 
volume_of_mimosa_lesions_lm <- lm(volume_of_mimosa_lesions ~ Diagnosis + osex + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_lm)

#gam lesion volumes; DiagnosisMS+Anxiety   6973.5     2014.7   3.461 0.000681 ***
volume_of_mimosa_lesions_gam <- gam(volume_of_mimosa_lesions ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_gam)
```

#### Anxiety dose
  - anxiety dose correlates with disease burden (volume of lesions)
```{r lesion_volumes_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != 3)

#dose; anxiety_dose     3028.07    1014.45   2.985 0.003026 **

  anxiety_dose_volume_of_mimosa_lesions_by_sex_lm <- lm(volume_of_mimosa_lesions ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_lm)

 #bilat anxiety dose + sex + age: anxiety_dose   3304.7      985.8   3.352 0.000886 ***

  anxiety_dose_volume_of_mimosa_lesions_by_sex_gam <- gam(volume_of_mimosa_lesions ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

### Uncinate **ACNP**

#### Demographics
```{r demo_uncinate}

###### all groups, no unclassified
make_demographics_table_ms_anxiety_simple(df_demo_and_fascicles %>% filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified"))

###### just healthy and sick
make_demographics_table_ms_anxiety_simple(df_demo_and_fascicles %>% filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds"))

```
#### Uncinate by diagnosis 
```{r uncinate_by_dx}

### Results: LM Mean UF_R & L, age, sex cov, DiagnosisMS+Anxiety   621.95     276.11   2.252   0.0256 * 
#define group
  df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol, mean_UF_vol) %>%
  ungroup()

#do lm UF*Diagnosis
 #UF_R_vol -> DiagnosisMS+Anxiety   793.14     361.94   2.191   0.0298 *   
disease_UF_R_vol_lm <- lm(UF_R_vol ~ Diagnosis + osex + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_UF_R_vol_lm)

 #UF_L_vol ->DiagnosisMS+Anxiety   450.75     261.20   1.726   0.0862 .    
disease_UF_L_vol_lm <- lm(UF_L_vol ~ Diagnosis + osex+PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_UF_L_vol_lm)

#bilat: DiagnosisMS+Anxiety   621.95     276.11   2.252   0.0256 *  
disease_mean_UF_vol_lm <- lm(mean_UF_vol ~ Diagnosis + osex+PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_mean_UF_vol_lm)

#gam
#UF_R_vol -> DiagnosisMS+Anxiety   795.60     366.58   2.170   0.0314 *   
disease_UF_R_vol_gam <- gam(UF_R_vol ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(disease_UF_R_vol_gam)

 #UF_L_vol -> DiagnosisMS+Anxiety   420.24     262.19   1.603    0.111   
disease_UF_L_vol_gam <- gam(UF_L_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(disease_UF_L_vol_gam)

#bilateral UF: DiagnosisMS+Anxiety   607.92     279.34   2.176   0.0309 *  
disease_mean_UF_vol_gam <- gam(mean_UF_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(disease_mean_UF_vol_gam)

#Fornix - Right side but not left
 #F_R_vol -> DiagnosisMS+Anxiety   860.50     368.16   2.337   0.0205 * 
#disease_F_R_vol_lm <- lm(F_R_vol ~ Diagnosis + osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
#summary(disease_F_R_vol_lm)

 #F_L_vol -> DiagnosisMS+Anxiety  329.724    181.609   1.816   0.0711 . 
#disease_F_L_vol_lm <- lm(F_L_vol ~ Diagnosis + osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
#summary(disease_F_L_vol_lm)
## net by sex 
```

#### Uncinate by Anxiety Dose 
```{r uncinateByAnxietyDose}

#results: Anxiety dose related to UF_R when covarying for linear effects of age and sex, just barely
# Additionally - taking the mean of both UF and UR, LM for bilat anxiety dose + sex + age: anxiety_dose most sig P= 0.0367 *

@@

#lm

>>>>

---
title: "MS Anxiety Longitudinal"
author: "Erica Baller"
date: "7/25/2024"
output:
  html_document: default
  pdf_document: default
---

@@

#findings

@@

#get list of scans with good Mimosa, n = 2263
mimosa_qc_gr_75 <- mimosa_qc_df %>%
  filter(score >= 75) %>%
  mutate(EXAM_DATA = as.character(EXAM_DATE)) %>%
  mutate(EMPI = as.character(EMPI)) %>%
  dplyr::select(EMPI, EXAM_DATE, score) %>%
  distinct(EMPI, EXAM_DATE, .keep_all = TRUE) %>%
  ungroup 
```

@@

#n=770, 1078 scans total
df_people_with_good_mimosa <- data_2023_pull %>% 
  distinct(EMPI, EXAM_DATE, .keep_all = TRUE) %>%
  group_by(EMPI) %>%
  arrange(EMPI, EXAM_DATE) %>%
  filter(row_number() <= 2) %>%
  left_join(mimosa_qc_gr_75, by = c("EMPI", "EXAM_DATE")) %>%
  filter(!is.na(score))

# Check the number of rows per EMPI and keep people with only 2, n scans=616, with n=308 subjects
df_people_with_2_rows <- df_people_with_good_mimosa %>%
  group_by(EMPI) %>%
  filter(n() == 2) %>%
  ungroup()

# View the result, n = 308, with 616 total scans
print(df_people_with_2_rows)
print(length(df_people_with_2_rows$EMPI))
print(length(unique(df_people_with_2_rows$EMPI)))
```

@@

 #n = 
df_demo_and_fascicles <- df_people_with_good_mimosa %>% 
  left_join(fascicle_proportions, by = c("EMPI", "EXAM_DATE")) %>%
  filter(!is.na(Has.MS.Provider)) %>% #remove people who don't have MS provider
  filter(!is.na(CNII_L)) %>%#this is a sample fascicle that should have a value. if it does not, it means something is usually wrong with whole person. Remove rows with NAs
  group_by(EMPI) %>% # again verify that this only includes people with two days
  filter(n() == 2) %>%
  mutate(exam_timepoint = ifelse(row_number() == 1, "t1", "t2"))

#check num unique EMPIs, n = 281
print(length(unique(df_demo_and_fascicles$EMPI)))

#check total n = 562
print(length(df_demo_and_fascicles$EMPI))

## Check to see the consistency within anxiety dose between t1 and t2
```{r consistency_of_anxiety}
# Create a summary to check for consistency, 127 people have two different anxiety doses
anxiety_dose_check <- df_demo_and_fascicles %>%
  group_by(EMPI) %>%
  summarize(
    unique_doses = n_distinct(anxiety_dose),
    doses = paste(unique(anxiety_dose), collapse = ", ")
  ) %>%
  ungroup()

# Filter to find EMPIs with more than one unique anxiety dose
inconsistent_doses <- anxiety_dose_check %>%
  filter(unique_doses > 1)

# Display the result
inconsistent_doses

#there are some, will take the higher value because if at some point you had disease, it should count as a lifetime history
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  group_by(EMPI) %>%
  mutate(lifetime_max_anxiety_dose = max(anxiety_dose))

#now check inconsistent doses
# Create a summary to check for consistency, 127 people have two different anxiety doses
anxiety_dose_check <- df_demo_and_fascicles %>%
  group_by(EMPI) %>%
  summarize(
    unique_doses = n_distinct(lifetime_max_anxiety_dose),
    doses = paste(unique(anxiety_dose), collapse = ", ")
  ) %>%
  ungroup()

# Filter to find EMPIs with more than one unique anxiety dose
inconsistent_doses <- anxiety_dose_check %>%
  filter(unique_doses > 1)

# Display the result - nothing there! It worked! Also checked this with my eyeballs
inconsistent_doses

@@

## Merge with data frame with good mimosas
```{r merge_lesion_volume_w_mimosa_df}
##########################################################
#             Merge lesion volumes with mimosa df        #
##########################################################
#merge whole data set with fascicle proportions, n = 809
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  left_join(lesion_volumes, by = c("EMPI", "EXAM_DATE"))

#check num unique EMPIs, n = 809
print(length(unique(df_demo_and_fascicles$EMPI)))

#check total n = 1618
print(length(df_demo_and_fascicles$EMPI))

@@

df_w_yeo_volume_info <- df_demo_and_fascicles %>%
  # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  dplyr::select(EMPI, EXAM_DATE, ends_with("_vol")) %>%
  ungroup()

@@

@@

@@

# Longitudinal Age*Disease

## Anxiety Diagnosis
```{r diagnosisXage}

#n healthy = 155 total, 122 unique, +minA = 131
#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(mean_UF_vol = (UF_L_vol + UF_R_vol)/2) %>%
  mutate(mean_UF_prop = (UF_L_prop + UF_R_prop)/2) %>%
#  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol, mean_UF_vol) %>%
  ungroup()

### diagnosis by age interaction?
#overall - NS
diagnosis_by_age <- lmer(proportion_volume_lost_per_total_network_size_sum ~ Diagnosis*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age)

### anxiety net
diagnosis_by_age_anx_net <- lmer(proportion_volume_lost_anxiety_net_by_network_size_sum_top_25 ~ Diagnosis*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_anx_net)

###nonanxiety net
diagnosis_by_age_nonanx_net <- lmer(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ Diagnosis*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_nonanx_net)

### UF_R
diagnosis_by_age_UF_R <- lmer(UF_R_vol ~ Diagnosis*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_UF_R)

### UF_L
diagnosis_by_age_UF_L <- lmer(UF_L_vol ~ Diagnosis*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_UF_L)

### mean UF
diagnosis_by_age_mean_UF <- lmer(mean_UF_vol ~ Diagnosis*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_mean_UF)
```

## Anxiety Dose
```{r diagnosisXage}
#n healthy = 135, +minA = 131, severeA = 60

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  mutate(EMPI = as.factor(EMPI)) %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(lifetime_max_anxiety_dose != -1) %>%
  mutate(anxiety_dose = as.numeric(lifetime_max_anxiety_dose)) %>%
  mutate(anxiety_dose_factor = ordered(anxiety_dose)) %>%
  #mean UF
  mutate(mean_UF_vol = (UF_L_vol + UF_R_vol)/2) %>%
  mutate(mean_UF_prop = (UF_L_prop + UF_R_prop)/2)

### diagnosis by age interaction?
#overall - NS
diagnosis_by_age <- lmer(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age)

diagnosis_by_age_factor <- lmer(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose_factor*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_factor)

### anxiety net
diagnosis_by_age_anx_net <- lmer(proportion_volume_lost_anxiety_net_by_network_size_sum_top_25 ~ anxiety_dose*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_anx_net)

###nonanxiety net
diagnosis_by_age_nonanx_net <- lmer(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_nonanx_net)

### UF_R
diagnosis_by_age_UF_R <- lmer(UF_R_vol ~ anxiety_dose*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_UF_R)

### UF_L
diagnosis_by_age_UF_L <- lmer(UF_L_vol ~ anxiety_dose*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_UF_L)

### mean UF
diagnosis_by_age_mean_UF <- lmer(mean_UF_vol ~ anxiety_dose*PAT_AGE_AT_EXAM + (1|EMPI), data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(diagnosis_by_age_mean_UF)
```

@@

### BEST SPLIT IS STRATIFYING ON ANXIETY MEDS!!!! Better than on variable that includes anxietyGroupVar(meds + anxiety diagnoses), or anxiety diagnosis alone
#make_demographics_table_ms_anxiety(data_frame = df_demo_and_fascicles)
#make_demographics_table_ms_anxiety(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$anxietyGroupVar != 0,])
#make_demographics_table_ms_anxiety(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$Has.Anxietydx.AND.AntianxietymedGroupVar != 0,])

@@

#COLLAPSE anxiety dx and antianxiety meds with those plus depression, and anxiety dx OR on anxiety meds into 1 group -->> THIS IS THE MONEY
### Overall Disease by Anxiety Dose
```{r group_anxietydx_meds_and_dep_comorbidity}
#this collapses anxiety group + meds with those with anxiety group + meds + with/without depression

@@

  #gam, anxiety_dose  
   anxietyGrouping_nonanxiety_net_gam<- gam(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_nonanxiety_net_gam)

  #volume lost in anxiety net by anxiety dose*sex: 
  anxietyGrouping_by_sex_nonanxiety_net_lm <- lm(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose*osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_by_sex_nonanxiety_net_lm)

  #gam, anxiety_dose         
  anxietyGrouping_by_sex_nonanxiety_net_gam <- gam(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose*osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_by_sex_nonanxiety_net_gam)

  #volume lost in anxiety net by anxiety dose + sex + PAT_AGE_AT_EXAM: 
  anxietyGrouping_cov_sex_nonanxiety_net_lm <- lm(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_cov_sex_nonanxiety_net_lm)

  #gam, anxiety_dose 
   anxietyGrouping_cov_sex_nonanxiety_net_gam <- gam(proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75 ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_cov_sex_nonanxiety_net_gam)

@@

###Check for polynomial relationships - Not finding them w healthy, anxietydx OR meds, anxiety AND meds
```{r polynomial_relationships}
  #check for polynomial relationship
#define group
df_demo_and_fascicles_no_unclass_anxiety_dose_factor <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != 3) %>% 
  mutate(anxiety_dose = as.factor(anxiety_dose))

@@

### Uncinate by Anxiety Dose
```{r uncinateByAnxietyDose}

#results: Anxiety dose related to UF_R when covarying for linear effects of age and sex, just barely

@@

########UF R - just looking at uncinate by anxiety dose; anxiety_dose      369.72     187.93   1.967   0.0499 *  *********** (LINEAR AGE)
  anxietyGrouping_UF_R_lm <- lm(UF_R_vol ~ anxiety_dose + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_lm)
